---
title: "Heat-pump operating frontier"
author: "CAKU"
date: "`r Sys.Date()`"
output: html_document
---

# Heat-pump operating frontier

This analysis combines the last 30 days of hourly weather observations for Kgs.
Lyngby with hourly electricity prices. It shows the average full consumer price
for each observed one-degree outdoor-temperature band and clock hour from 07:00
to 23:00. Dark blue cells are the lowest-price combinations; white cells are
the highest-price combinations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(lubridate)
source("R/data.R")
source("R/tariffs.R")
source("R/weather.R")
source("R/plot-theme.R")
```

```{r data, include=FALSE, cache=TRUE}
period_start <- Sys.Date() - 30
period_end <- Sys.Date() - 1

weather <- fetch_historical_weather(period_start, period_end) %>%
  mutate(HourKey = as.numeric(as.POSIXct(HourDK, tz = "UTC")) %/% 3600) %>%
  select(-HourDK)

prices <- bind_rows(
  fetch_elspot_prices("DK1", start = format(period_start), end = format(Sys.Date()),
                      require_future = FALSE),
  fetch_elspot_prices("DK2", start = format(period_start), end = format(Sys.Date()),
                      require_future = FALSE)
) %>%
  apply_consumer_price_components(.) %>%
  rename(ConsumerPriceDKK = SpotPriceDKK_Total) %>%
  mutate(HourKey = as.numeric(as.POSIXct(HourDK, tz = "UTC")) %/% 3600)

match_weather_prices <- function(weather, prices) {
  candidate_offsets <- c(0, -1, 1, -2, 2, -24, 24)
  candidate_matches <- lapply(candidate_offsets, function(offset) {
    weather %>%
      mutate(HourKey = HourKey + offset) %>%
      inner_join(prices, by = "HourKey") %>%
      arrange(PriceArea, HourDK)
  })
  candidate_matches[[which.max(vapply(candidate_matches, nrow, integer(1)))]]
}

heatpump_data <- match_weather_prices(weather, prices)
has_target_hours <- function(data) {
  nrow(data) > 0 && any(lubridate::hour(data$HourDK) %in% 7:23)
}
if (!has_target_hours(heatpump_data)) {
  weather <- fetch_historical_weather_open_meteo(period_start, period_end) %>%
    mutate(HourKey = as.numeric(as.POSIXct(HourDK, tz = "UTC")) %/% 3600) %>%
    select(-HourDK)
  heatpump_data <- match_weather_prices(weather, prices)
}
if (!nrow(heatpump_data)) {
  stop("No weather/price observations could be matched after timezone alignment.",
       call. = FALSE)
}

frontiers <- heatpump_data %>%
  group_by(PriceArea) %>%
  group_modify(~ pareto_frontier(.x)) %>%
  ungroup()

heatmap_data <- heatpump_data %>%
  mutate(
    TemperatureBand = floor(TemperatureC),
    HourBand = lubridate::hour(lubridate::with_tz(
      as.POSIXct(HourKey * 3600, origin = "1970-01-01", tz = "UTC"),
      tzone = weather_config$timezone
    ))
  ) %>%
  filter(HourBand >= 7, HourBand <= 23) %>%
  group_by(PriceArea, TemperatureBand, HourBand) %>%
  summarise(
    ConsumerPriceDKK = mean(ConsumerPriceDKK, na.rm = TRUE),
    Observations = n(),
    .groups = "drop"
  )

if (!nrow(heatmap_data)) {
  stop("No matched weather/price observations were available between 07:00 and 23:00.",
       call. = FALSE)
}
```

## Heatmap: temperature, hour and electricity price

Each tile represents the mean full consumer price for a one-degree temperature
band and one clock hour. This is a decision-support view, not a manufacturer
COP curve: actual COP also depends on flow temperature, system design,
defrosting and the specific heat pump.

```{r heatmap, fig.width=10, fig.height=7}
ggplot(heatmap_data,
       aes(x = TemperatureBand, y = HourBand, fill = ConsumerPriceDKK)) +
  geom_tile(colour = "#173238", linewidth = 0.25) +
  facet_wrap(~ PriceArea, ncol = 1) +
  scale_fill_gradient(
    low = "#062b5c", high = "#ffffff", na.value = "#173238",
    name = "DKK/kWh"
  ) +
  scale_y_continuous(breaks = 7:23, limits = c(6.5, 23.5), expand = c(0, 0)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 12), expand = c(0, 0)) +
  labs(
    title = "Heat-pump electricity-price heatmap, last 30 days",
    subtitle = paste(period_start, "to", period_end,
                     "· hourly bands · full Andel FlexEnergi consumer price"),
    x = "Outdoor temperature band (°C)",
    y = "Hour of day",
    caption = "Dark blue = lowest observed mean price; white = highest."
  ) +
  plot_theme()
```

```{r summary}
heatmap_data %>%
  group_by(PriceArea) %>%
  summarise(
    `Temperature bands` = n_distinct(TemperatureBand),
    `Hour bands` = n_distinct(HourBand),
    `Lowest mean price` = scales::number(min(ConsumerPriceDKK), accuracy = 0.01),
    `Highest mean price` = scales::number(max(ConsumerPriceDKK), accuracy = 0.01),
    .groups = "drop"
  ) %>% knitr::kable()
```

The OpenWeatherMap historical endpoint must be enabled for the
`OPENWEATHERMAP_APIKEY` used by the build. If that key or historical access is
missing, the report stops rather than presenting an incomplete frontier.
