---
title: "Heat-pump operating frontier"
author: "CAKU"
date: "`r Sys.Date()`"
output: html_document
---

# Heat-pump operating frontier

This analysis combines the last 30 days of hourly weather observations for Kgs.
Lyngby with hourly electricity prices. It highlights hours that offer a useful
trade-off between a warmer outdoor temperature (a transparent COP proxy) and a
lower full consumer electricity price.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(lubridate)
source("R/data.R")
source("R/tariffs.R")
source("R/weather.R")
source("R/plot-theme.R")
```

```{r data, include=FALSE, cache=TRUE}
period_start <- Sys.Date() - 30
period_end <- Sys.Date() - 1

weather <- fetch_historical_weather(period_start, period_end)

prices <- bind_rows(
  fetch_elspot_prices("DK1", start = format(period_start), end = format(Sys.Date()),
                      require_future = FALSE),
  fetch_elspot_prices("DK2", start = format(period_start), end = format(Sys.Date()),
                      require_future = FALSE)
) %>%
  apply_consumer_price_components(.) %>%
  rename(ConsumerPriceDKK = SpotPriceDKK_Total)

heatpump_data <- prices %>%
  inner_join(weather, by = "HourDK") %>%
  arrange(PriceArea, HourDK)

frontiers <- heatpump_data %>%
  group_by(PriceArea) %>%
  group_modify(~ pareto_frontier(.x)) %>%
  ungroup()
```

## Temperature versus full electricity price

Each point is one matched hour. The frontier contains hours for which no other
observed hour is both warmer and cheaper. This is a decision-support view, not
a manufacturer COP curve: actual COP also depends on flow temperature, system
design, defrosting and the specific heat pump.

```{r frontier, fig.width=10, fig.height=6}
ggplot(heatpump_data,
       aes(x = TemperatureC, y = ConsumerPriceDKK, colour = PriceArea)) +
  geom_point(alpha = 0.28, size = 1.7) +
  geom_line(data = frontiers,
            aes(group = PriceArea), linewidth = 1.1, inherit.aes = TRUE) +
  geom_point(data = frontiers, size = 2.2, inherit.aes = TRUE) +
  labs(
    title = "Heat-pump operating frontier, last 30 days",
    subtitle = paste(period_start, "to", period_end,
                     "· full Andel FlexEnergi consumer price"),
    x = "Outdoor temperature (°C; higher generally improves COP)",
    y = "Electricity price (DKK/kWh, all components)",
    colour = "Price area"
  ) +
  plot_theme()
```

```{r summary}
frontiers %>%
  group_by(PriceArea) %>%
  summarise(
    `Matched hours` = n(),
    `Cheapest frontier price` = scales::number(min(ConsumerPriceDKK), accuracy = 0.01),
    `Warmest frontier temperature` = scales::number(max(TemperatureC), accuracy = 0.1),
    .groups = "drop"
  ) %>%
  knitr::kable()
```

The OpenWeatherMap historical endpoint must be enabled for the
`OPENWEATHERMAP_APIKEY` used by the build. If that key or historical access is
missing, the report stops rather than presenting an incomplete frontier.
